Parsing Postfix Log Files

### 2009-04-23 doc/logparser-architecture.tex
Changes based on comments by Ariane.

\begin{table}[thbp]
    \caption{Similarities between ATN and this architecture}
    \empty{}\label{Similarities between ATN and this architecture}
    \begin{tabular}[]{lll}
        \tabletopline{}%
        \acronym{ATN}   & Architecture  & Similarity                  \\
        \tablemiddleline{}%
        Networks        & Parser        & Determines the sequence
                                          of transitions              \\
                        &               & or actions that
                                          constitutes a valid input.  \\
        Transitions     & Actions       & Assemble data and
                                          impose conditions the       \\
                        &               & input must meet to be
                                          accepted as valid.          \\
        Abbreviations   & Rules         & Responsible for
                                          classifying input.          \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

The architecture can be thought of as implementing transduction: it takes
data in one form and transforms it to another form; \parsername{}
transforms log files to a database.

@@

The framework manages the parsing process and provides support functions
for the actions, freeing the programmers writing actions to concentrate on
writing productive code.  It links actions and rules, allowing either to be
improved independently of the other, and allows new rules to be written
without needing changes to the source code of a parser.  The framework is
the core of the architecture and is deliberately quite simple: the rules
deal with the variation in inputs, and the actions deal with the
intricacies and complications encountered during parsing.  The function
that finds the rule that recognises the input and invokes the requested
action can be expressed in pseudo-code as:

\begin{verbatim}
INPUT:
for each input {
    for each rule defined by the user {
        if this rule recognises the input {
            invoke the action specified by the rule
            next INPUT
        }
    }
    warn the user that the input was not recognised
}
\end{verbatim}

@@

    \item [Classify the input]  The pseudo-code above shows how rules are
        successively tried until one is found that recognises the current
        input.  That pseudo-code is very simple: there may be efficiency
        concerns (\sectionref{parser efficiency}), rule conditions
        (\sectionref{attaching conditions to rules}), or rule priorities
        (\sectionref{rules in architecture}) that complicate the process.

>>>>

\begin{table}[thbp]
    \caption{Similarities between ATN and this architecture}
    \empty{}\label{Similarities between ATN and this architecture}
    \centering{}
    \begin{tabular}[]{lll}
        \tabletopline{}%
        \acronym{ATN}   & Architecture  & Similarity                  \\
        \tablemiddleline{}%
        Networks        & Framework     & Determines the sequence
                                          of transitions              \\
                        &               & or actions that
                                          constitutes a valid input.  \\
        Transitions     & Actions       & Assemble data and
                                          impose conditions the       \\
                        &               & input must meet to be
                                          accepted as valid.          \\
        Abbreviations   & Rules         & Responsible for
                                          classifying input.          \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

The architecture can be thought of as implementing transduction: it takes
data in one form and transforms it to another form; \parsername{}
transforms log files to a \acronym{SQL} database.

@@

The framework manages the parsing process and provides support functions
for the actions, freeing the programmers writing actions to concentrate on
writing productive code.  It links actions and rules, allowing either to be
improved independently of the other, and allows new rules to be written
without needing changes to the source code of a parser.  The framework is
the core of the architecture and is deliberately quite simple: the rules
deal with the variation in inputs, and the actions deal with the
intricacies and complications encountered during parsing.  The function
that finds the rule that recognises the input and invokes the requested
action can be expressed in pseudo-code as shown in \figureref{Pseudo-code
describing how inputs are recognised}.

\begin{figure}[thbp]
    \caption{Pseudo-code describing how inputs are recognised}
    \empty{}\label{Pseudo-code describing how inputs are recognised}
\begin{verbatim}
INPUT:
for each input {
    for each rule defined by the user {
        if this rule recognises the input {
            invoke the action specified by the rule
            next INPUT
        }
    }
    warn the user that the input was not recognised
}
\end{verbatim}
\end{figure}

@@

    \item [Classify the input]  The pseudo-code in \figureref{Pseudo-code
        describing how inputs are recognised} shows how rules are
        successively tried until one is found that recognises the current
        input.  That pseudo-code is very simple: there may be efficiency
        concerns (\sectionref{parser efficiency}), rule conditions
        (\sectionref{attaching conditions to rules}), or rule priorities
        (\sectionref{rules in architecture}) that complicate the process.

### 2009-04-23 doc/logparser-other-parsers-review.tex
Changes based on comments by Ariane.

The ten parsers have been reviewed and compared with \parsername{}, this
project's parser, to show how much effort would have been required to use
one of those parsers to fulfil the aims and requirements of this project.
It is important to compare and contrast newly developed algorithms and
parsers against those already available, to accurately judge what
improvements, if any, are delivered by the newcomers.

@@

    \item None of the reviewed parsers perform the kind of advanced parsing
        required for this project, or deal with the complications described
        in \sectionref{complications}.

@@

Each of the reviewed parsers was tested with the \numberOFlogFILES{} test
log files described in \sectionref{parser efficiency}; \tableref{Summary of
parsers' features} summarises the results of this review.

The data extracted by \parsername{} is documented in
\sectionref{connections table} and \sectionref{results table}; for
convenience that list is repeated here: client and server IP address and
hostname, HELO hostname, queueid, start time, end time, \acronym{SMTP}
code, enhanced status code, sender, recipient, size, message-id, delay, and
delays.

@@

\acronym{CSV} is a very simple format where each record is stored in a
single line, with fields separated by a comma or other punctuation symbol.
Problems with \acronym{CSV} files include the need to escape separators in
the data stored, providing multiple values for a field (e.g.\ multiple
recipients), and adding new fields.  \acronym{CSV} files do not have a
standard mechanism to document the fields or the separator, unlike
\acronym{SQL} where every database includes a schema naming the fields and
the type of data they store (e.g.\ integer, text, timestamp).  The
\acronym{CSV} record format used by \acronym{LMA} is not documented, but
the output file contains a comment giving the format:\newline{} \texttt{\#
Timestamp|Nome Client|IP Client|IP Server|From|To|Status|Size}
\newline{}\acronym{LMA} treats \acronym{CSV} lines starting with
\texttt{\#} as comments, but not all \acronym{CSV} parsers will.

@@

The schema used with the MySQL database is undocumented, but at least it is
possible to discover the schema with an existing \acronym{SQL} database,
unlike with Berkeley DB\@; all \acronym{SQL} databases embed the schema
into the database and provide commands for displaying it.  Berkeley DB does
not embed a schema in its files, because it does not need one, and
embedding one would not provide any benefit; it only provides \textbf{(key,
value)} pairs, so any additional structuring of the data is imposed by the
application, and it is the responsibility of the author of the application
to document this structure.  MySQL support was not tested because the
schema is not documented, so the required tables could not be created.

@@

\acronym{LMA} does provide some simple reports: IP-STORY, FROM-STORY,
DAILY-EMAIL and DAILY-REJECT\@.  These reports search the Berkeley DB files
for matching records: the first three produce \acronym{CSV} lines for the
specified client IP address, sender address, or date respectively.
DAILY-REJECT initially failed with an error message from the Perl
interpreter;\footnote{The error messages were: \newline{}\texttt{Undefined
subroutine \&main::LIST called at queryDB.pl line
372.}\newline{}\texttt{Undefined subroutine \&main::EXTRACT\_FROM\_DB
called at queryDB.pl line 379.}} after correcting the errors in the code it
worked, producing \acronym{CSV} lines for the specified day where the
\acronym{SMTP} code signifies a rejection.  All of these reports are
trivially simple to produce from the \acronym{CSV} file using the standard
Unix tool \texttt{awk}\glsadd{awk}; the most complicated, DAILY-REJECT, is
merely:

@@

The source code is only available in an encrypted form, to support people
who wish to use Sawmill on operating systems or machine architectures the
company do not provide executables for.  Sawmill is quite expensive,
requiring a \euros{100} + VAT licence per report, with discounts available
when buying multiple licences (correct as of 2008/11/23); in contrast,
\parsername{} is free to use and the code is freely available.  Sawmill is
supplied with thorough and well written documentation; everything the
author searched for was documented, except the MySQL database schema and
some details of the data extracted by the parser.  A commercial version of
MySQL is required because of MySQL licensing restrictions, but Sawmill's
documentation explains why and includes instructions on how to compile
Sawmill so that it can use a non-commercial version of MySQL (this was not
attempted during the review process).

@@

Splunk aims to index all of an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface provides search functionality, generating statistics and
graphs in real time, a facility not provided by \parsername{}.  Splunk
allows quite complicated searches, based on the fields it extracts
(described later) or the full text of the log line, though it is not
possible to search on partial words.  Search results are not available for
use with other tools.  Searches can be saved for reuse; saved searches can
be run periodically and the results mailed to a recipient or sent to an
external program for further processing.  The author was unable to save
searches, though that may have been because of limitations in the free
version.  The web interface is optimised for interactive use rather than
automated queries, and it does not appear to be possible to write
independent tools to utilise the Splunk database, whereas \parsername{}
provides the database and leaves it to the user to utilise it without limit
or restriction.  Some additional Postfix reports are supposedly available
on \urlLastChecked{http://www.splunkbase.com/}{2009/04/21}, but the author
was unable to find any Postfix reports, or indeed reports for any other log
file types: every category was empty, even those that the web site claimed
had numerous reports available.  Many types of graphs can be generated,
though all except the bubble and heatmap graphs are variations of a bar or
pie chart.  It is trivial to drill down through the graphs to select a
portion of the data, e.g.\ select the hour with the largest number of
events, then select a particular host, and finally a specific sender
address.  All searches performed using the indexed data returned reasonable
results.  The full power of \acronym{SQL} is available when searching the
data extracted by \parsername{}, allowing the user to search on arbitrarily
complicated conditions.

@@

When parsing Postfix log files Splunk parses the standard
syslog\glsadd{syslog} fields at the beginning of the log line, and extracts
any \texttt{key=value} pairs occurring after the syslog prologue: to and
from addresses, HELO hostname, and protocol (\acronym{SMTP},
\acronym{LMTP}, or \acronym{ESMTP}).  \parsername{} extracts noticeably
more data (client and server IP address and hostname, queueid, start and
end times, timestamps for each log line, \acronym{SMTP} and enhanced status
codes, delivery delays, and message-id), though it does not make the full
text of the log line available; this could be trivially added if desired,
but would greatly increase the size of the resulting database.

@@

\begin{table}[thbp]
    \caption{Summary of reviewed parsers' features}
    \empty{}\label{Summary of parsers' features}
    \begin{tabular}{llllll}
        \tabletopline{}%
        Parser          & Parsed test   & Data              & Custom            & Documentation  & Source       \\
                        & log files?    & store?            & reports?          & quality?       & code?        \\
        \tablemiddleline{}%
        \acronym{LMA}       & No            & Yes               & No                & Poor           & Yes          \\
        Pflogsumm       & Yes           & No                & Partial \dag{}    & Good           & Yes          \\
        Sawmill         & Yes           & Yes               & Searches          & Excellent      & \nialpha{}   \\
        Splunk          & Yes           & Yes               & Searches          & Abundant       & No           \\
                        &               &                   & \& reports        & but poor       &              \\
        Isoqlog         & No            & Yes               & No                & No English     & Yes          \\
                        &               &                   &                   & documentation  &              \\
        AWStats         & Partially     & Yes               & Partial \dag{}    & Good           & Yes          \\
        Anteater        & No            & No                & No                & None           & Yes          \\
        YAALA           & No            & Yes \ddag{}       & Searches          & Poor           & Yes          \\
                        &               &                   & \& reports        &                &              \\
        Lire            & Yes           & Yes               & Yes               & Reasonable     & Yes          \\
        Logrep          & No            & Yes               & No                & None           & Yes          \\
        \parsername{}   & Yes           & Yes \nibeta{}     & No \nichi{}       & \niepsilon{}   & Yes          \\
        \tablebottomline{}%
    \end{tabular}

@@

        \item [\nialpha{}] Sawmill's source code is available in an
            encrypted form, so that customers can compile it on platforms
            that pre-compiled binaries are not available for.

>>>>

The ten parsers have been reviewed and compared with \parsername{}, this
project's parser, to show how much effort would have been required to use
one of those parsers to fulfil the aims and requirements of this project.
It is important to compare and contrast newly developed algorithms and
parsers against those already available, to accurately judge what
improvements, if any, are delivered by the newcomers.  \Tableref{Summary of
parsers' features} summarises the results of this review.

@@

    \item None of the reviewed parsers perform the kind of advanced parsing
        required for this project, or deal with the complications described
        in \sectionref{complications}.  Some correlate log lines by
        queueid, but none correlate log lines by \acronym{pid}
        (\sectionref{queueid vs pid}).

@@

Each of the reviewed parsers was tested with the three months
(\numberOFlogFILES{} days) of contiguous log files described in
\sectionref{parser efficiency}.

The data extracted by \parsername{} is documented in
\sectionref{connections table} and \sectionref{results table}; for
convenience that list is repeated here: client and server IP address and
hostname, HELO hostname, queueid, start time, end time, \acronym{SMTP}
code, enhanced status code, sender, recipient, size, message-id, delay, and
delays.  Note: \parsername{} correlates log lines by both queueid and
\acronym{pid} (\sectionref{queueid vs pid}), and it stores each mail's
queueid, but it does not store each connection's \acronyms{pid}, because a
use has not been identified for that data.

@@

\acronym{CSV} is a very simple format where each record is stored in a
single line, with fields separated by a comma or other punctuation symbol.
Problems with \acronym{CSV} files include the need to escape separators in
the data stored, providing multiple values for a field (e.g.\ multiple
recipients), and adding new fields.  \acronym{CSV} files do not have a
standard mechanism to document the fields or the separator, unlike
\acronym{SQL} where every database includes a schema naming the fields and
the type of data they store (e.g.\ integer, text, timestamp).  The
\acronym{CSV} record format used by \acronym{LMA} is not documented, but
the output file contains a comment in Italian giving the format:\newline{}
\texttt{\# Timestamp|Nome Client|IP Client|IP Server|From|To|Status|Size}
\newline{}\acronym{LMA} treats \acronym{CSV} lines starting with
\texttt{\#} as comments, but not all \acronym{CSV} parsers will.

@@

The schema used with the MySQL database is undocumented, but at least it is
possible to discover the schema with an existing \acronym{SQL} database,
unlike with Berkeley DB\@; all \acronym{SQL} databases embed the schema
into the database and provide commands for displaying it.  Berkeley DB does
not embed a schema in its files, because it does not need one, and
embedding one would not provide any benefit; it only provides \textbf{(key,
value)} pairs, so any additional structuring of the data is imposed by the
application, and it is the responsibility of the author of the application
to document this structure.  MySQL support was not tested because the
schema is not documented, and no existing database was available to extract
the schema from, so the required tables could not be created.

@@

\acronym{LMA} does provide some simple reports: IP-STORY, FROM-STORY,
DAILY-EMAIL and DAILY-REJECT\@.  These reports search the Berkeley DB files
for matching records: the first three produce \acronym{CSV} lines for the
specified client IP address, sender address, or date respectively.
DAILY-REJECT initially failed with an error message from the Perl
interpreter;\footnote{The error messages were: \newline{}\texttt{Undefined
subroutine \&main::LIST called at queryDB.pl line
372.}\newline{}\texttt{Undefined subroutine \&main::EXTRACT\_FROM\_DB
called at queryDB.pl line 379.}} after correcting the errors in the code it
worked, producing \acronym{CSV} lines for the specified day where the
\acronym{SMTP} code signifies a rejection.  All of these reports are
extremely simple to produce from the \acronym{CSV} file using the standard
Unix tool \texttt{awk}\glsadd{awk}; the most complicated, DAILY-REJECT, is
merely:

@@

The source code is only available in an encrypted form, to support people
who wish to use Sawmill on operating systems or machine architectures the
company do not provide executables for.  Sawmill is quite expensive,
requiring a \euros{100} + VAT licence per report, with discounts available
when buying multiple licences (correct as of 2008/11/23); in contrast,
\parsername{} is free to use and the code is freely available.  Sawmill is
supplied with thorough and well written documentation; everything the
author looked for was documented, except for the MySQL database schema and
some details of the data extracted by the parser, e.g.\ what the virus
field stores.  A commercial version of MySQL is required because of MySQL
licensing restrictions, but Sawmill's documentation explains why and
includes instructions on how to compile Sawmill so that it can use a
non-commercial version of MySQL (this was not attempted during the review
process).

@@

Splunk aims to index all of an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface provides search functionality, generating statistics and
graphs in real time, a facility not provided by \parsername{}.  Splunk
allows quite complicated searches, based on the fields it extracts
(described later) or the full text of the log line, though it is not
possible to search on partial words.  Searches can be saved for reuse;
saved searches can be run periodically and the results mailed to a
recipient or sent to an external program for further processing.  The
author was unable to save searches, possibly because of limitations in the
free version, and so was unable to examine the format of the data.  The web
interface is optimised for interactive use rather than automated queries,
and it does not appear to be possible to write independent tools to utilise
the Splunk database, whereas \parsername{} provides the database and leaves
it to the user to utilise it without limit or restriction.  Some additional
Postfix reports are supposedly available on
\urlLastChecked{http://www.splunkbase.com/}{2009/04/21}, but the author was
unable to find any Postfix reports, or indeed reports for any other log
file types: every category was empty, even those that the web site claimed
had numerous reports available.  Many types of graphs can be generated,
though all except the bubble and heatmap graphs are variations of a bar or
pie chart.  Drilling down through the graphs to select a portion of the
data is simple and intuitive, e.g.\ select the hour with the largest number
of events, then select a particular host, and finally a specific sender
address.  All searches performed using the indexed data returned reasonable
results.  The full power of \acronym{SQL} is available when searching the
data extracted by \parsername{}, allowing the user to search on arbitrarily
complicated conditions.

@@

When parsing Postfix log files Splunk parses the standard
syslog\glsadd{syslog} fields at the beginning of the log line, and extracts
any \texttt{key=value} pairs occurring after the syslog prologue: to and
from addresses, HELO hostname, and protocol (\acronym{SMTP},
\acronym{LMTP}, or \acronym{ESMTP}).  \parsername{} extracts noticeably
more data (client and server IP address and hostname, queueid, start and
end times, timestamps for each log line, \acronym{SMTP} and enhanced status
codes, delivery delays, and message-id), though it does not make the full
text of the log line available; a few minutes work could add this to
\parsername{} if desired, but it would greatly increase the size of the
database.

@@

\begin{table}[thbp]
    \caption{Summary of reviewed parsers' features}
    \empty{}\label{Summary of parsers' features}
    \begin{tabular}{llllll}
        \tabletopline{}%
        Parser          & Parsed test   & Data              & Custom            & Documentation  & Source           \\
                        & log files?    & store?            & reports?          & quality?       & code?            \\
        \tablemiddleline{}%
        \acronym{LMA}   & No            & Yes               & No                & Poor           & Yes              \\
        Pflogsumm       & Yes           & No                & Partial \dag{}    & Good           & Yes              \\
        Sawmill         & Yes           & Yes               & Searches          & Excellent      & Yes \nialpha{}   \\
        Splunk          & Yes           & Yes               & Searches          & Abundant       & No               \\
                        &               &                   & \& reports        & but poor       &                  \\
        Isoqlog         & No            & Yes               & No                & No English     & Yes              \\
                        &               &                   &                   & documentation  &                  \\
        AWStats         & Partially     & Yes               & Partial \dag{}    & Good           & Yes              \\
        Anteater        & No            & No                & No                & None           & Yes              \\
        YAALA           & No            & Yes \ddag{}       & Searches          & Poor           & Yes              \\
                        &               &                   & \& reports        &                &                  \\
        Lire            & Yes           & Yes               & Yes               & Poor           & Yes              \\
        Logrep          & No            & Yes               & No                & None           & Yes              \\
        \parsername{}   & Yes           & Yes \nibeta{}     & No \nichi{}       & \niepsilon{}   & Yes              \\
        \tablebottomline{}%
    \end{tabular}

@@

        \item [\nialpha{}] Sawmill's source code is available, but in an
            encrypted or obfuscated form.

### 2009-04-23 doc/logparser-results.tex
Changes based on comments by Ariane.

When generating the timing data used in this section, \numberOFlogFILES{}
log files were used, each containing one day's log lines, totalling
\input{build/include-total-size-of-93-log-files}, \numberOFlogLINEShuman{}
log lines.  Each log file was parsed 10 times, and the mean parsing time
calculated.  The computer used for test runs was a Dell Optiplex 745, shown
in \tableref{Details of the computer used to generate statistics}; it was
dedicated to the task of gathering statistics from test runs, and did not
run any other programs while test runs were in progress.  Saving results to
the database was disabled for the test runs, because that dominates the run
time of the parser, and the tests are aimed at measuring the speed of
\parsername{} rather than the speed of the database and the disks it is
stored on.  Parsing all \numberOFlogFILES{} log files in one run without
saving results to the database took
\input{build/include-full-run-duration}, with mean throughput of
\input{build/include-full-run-throughput}.  In contrast, when saving
results to the database, parsing all \numberOFlogFILES{} log files took
\input{build/include-insert-results-duration}, with mean throughput of
\input{build/include-insert-results-throughput} --- parsing takes up only
\input{build/include-skip-inserting-as-percentage-of-inserting} of the
execution time when saving results to the database.

>>>>

When generating the timing data used in this section, \numberOFlogFILES{}
log files were used, each containing one day's log lines, totalling three
months of contiguous log files,
\input{build/include-total-size-of-93-log-files}, \numberOFlogLINEShuman{}
log lines.  Each log file was parsed 10 times, and the mean parsing time
calculated.  The computer used for test runs was a Dell Optiplex 745, shown
in \tableref{Details of the computer used to generate statistics}; it was
dedicated to the task of gathering statistics from test runs, and did not
run any other programs while test runs were in progress.  Saving results to
the database was disabled for the test runs, because that dominates the run
time of the parser, and the tests are aimed at measuring the speed of
\parsername{} rather than the speed of the database and the disks it is
stored on.  Parsing all \numberOFlogFILES{} log files in one run without
saving results to the database took
\input{build/include-full-run-duration}, with mean throughput of
\input{build/include-full-run-throughput}.  In contrast, when saving
results to the database, parsing all \numberOFlogFILES{} log files took
\input{build/include-insert-results-duration}, with mean throughput of
\input{build/include-insert-results-throughput} --- parsing takes up only
\input{build/include-skip-inserting-as-percentage-of-inserting} of the
execution time when saving results to the database.
